Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-08 paper.tex
small edits

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \system,
does three things:
it finds programs that solve each of the tasks;
it creates its own DSL by discovering and reusing
domain-specific subroutines;
and it learns a neural network that
efficiently searches for programs in the DSL.
The purpose of the DSL
is to distill the commonalities across the programs that solve the tasks,
and it can be used to solve other related program learning problems.

@@

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', which draws samples from a recognition model (our $q$);
and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
The wake cycle can be thought of as a kind of active problem solving,
guided by the current generative and recognition models.
A sleep cycle means updating these models so that the DSL and  recognition model
to more closely match the distribution of tasks.
This wake/sleep style of inference is the reason why we call our approach \system.

@@

\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
%\\\noindent\textbf{Definition.}
A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
one number for each DSL primitive $e:\tau\in \mathcal{D}$, written $\theta_e$,
and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{c}
    String editing primitives\\
    \begin{tabular}{l}
    \toprule
    \code{0}, \code{+1},\code{-1}\\
    \code{split}, \code{join}, \code{map}, \code{length}, \code{slice}\\
    \code{append}, \code{chr->str}, \code{trim}, \code{nth}\\
    \code{lowercase}, \code{uppercase}, \code{capitalize}\\
    \code{""},    \code{","},     \code{" "},    \code{"<"}, \code{">"}  
  \\\bottomrule
    \end{tabular}\\
\\    Subset of learned DSL\\
    \begin{tabular}{l}
      \toprule
      $f_0($\code{c}$) = $\code{(+ (chr->str c))}\\
      $f_1($\code{i,c,s}$) = $      \code{(+ (nth i (split c s))}\\\hspace{2.55cm}\code{ (chr->str c))}\\
      $f_2($\code{s,c}$) = $\code{((}$f_0$\code{ c) (}$f_1$\code{ (1+ 0) c s))} \\
      \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
      $f_3($\code{c,s}$) = $\code{(nth (1+ 0) (split c (nth 0 s)))}\\
      \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
    %% $f_0($\code{i,c,s}$) = $    \code{(+ (index i (split c s)) (chr->str c))}\\
    %% $f_1 = $\code{(join "")}
  \\\bottomrule
    \end{tabular}
  \end{tabular}
  \caption{Top: provided string editing primitives. Bottom: some induced DSL primitives.}\label{textPrimitives}
  \end{table}
%% \begin{figure}

>>>>

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \system,
does three things:
it finds programs that solve each of the tasks;
it creates its own DSL by discovering and reusing
domain-specific subroutines;
and it learns a neural network that
efficiently searches for programs in the DSL.
The purpose of the DSL
is to distill the commonalities across the programs that solve the tasks,
letting the agent solve other related program learning problems.

@@

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', which draws samples from a recognition model (our $q$);
and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
The wake cycle can be thought of as a kind of active problem solving,
guided by the current generative and recognition models.
A sleep cycle means updating these models so that the DSL and  recognition model
more closely match the distribution of tasks.
This wake/sleep style of inference is the reason why we call our approach \system.

@@

\noindent\textbf{Definition: $(\mathcal{D},\theta)$.}
A DSL $\mathcal{D}$ is a set of typed $\lambda$-calculus expressions.
%\\\noindent\textbf{Definition.}
A weight vector $\theta$ for a DSL $\mathcal{D}$ is a vector of $|\mathcal{D}| + 1$ real numbers:
one number for each DSL primitive $e\in \mathcal{D}$, written $\theta_e$,
and a weight controlling the probability of a variable occurring in a program, $\theta_{\text{var}}$.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{c}
    String editing primitives\\
    \begin{tabular}{l}
    \toprule
    \code{0}, \code{+1},\code{-1},     \code{""},    \code{","},     \code{" "},    \code{"<"}, \code{">"}  \\
    \code{split}, \code{join}, \code{map}, 
    \code{append}, \code{chr->str}, \code{trim}, \code{nth}\\
    \code{length}, \code{slice},    \code{lower}, \code{upper}, \code{capitalize}
  \\\bottomrule
    \end{tabular}\\
\\    Subset of learned DSL\\
    \begin{tabular}{l}
      \toprule
      $f_0($\code{c}$) = $\code{(+ (chr->str c))}\\
      $f_1($\code{i,c,s}$) = $      \code{(+ (nth i (split c s))}\\\hspace{2.55cm}\code{ (chr->str c))}\\
      $f_2($\code{s,c}$) = $\code{((}$f_0$\code{ c) (}$f_1$\code{ (1+ 0) c s))} \\
      \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
      $f_3($\code{c,s}$) = $\code{(nth (1+ 0) (split c (nth 0 s)))}\\
      \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
    %% $f_0($\code{i,c,s}$) = $    \code{(+ (index i (split c s)) (chr->str c))}\\
    %% $f_1 = $\code{(join "")}
  \\\bottomrule
    \end{tabular}
  \end{tabular}
  \caption{Top: provided string editing primitives. Bottom: some induced DSL primitives.}\label{textPrimitives}
  \end{table}
%% \begin{figure}
